---
title: "doubletFinderTrial"
author: "Drew"
date: "2024-06-26"
output: html_document
---

```{r}

```


```{r}
library(Seurat)
library(patchwork)
library(dplyr)
library(presto)
library(gridExtra)
library(grid)

library(Matrix)
library(fields)
library(KernSmooth)
library(ROCR)
library(parallel)
library(DoubletFinder)
```


### Reading in the data 

```{r}

datDir <- "~/Uni/research project/R_researchProject/Data/Filtered2Markers"
list.files(datDir)

rawDataTu <- Read10X(data.dir = file.path(datDir, "Tumor"))
rawData9 <- Read10X(data.dir = file.path(datDir, "9weeks"))
rawData6 <- Read10X(data.dir = file.path(datDir, "6weeks"))
rawData3 <- Read10X(data.dir = file.path(datDir, "3weeks"))

datTu <- CreateSeuratObject(counts = rawDataTu, project = "tumor", min.cells = 3, min.features = 200)
dat9 <- CreateSeuratObject(counts = rawData9, project = "week9", min.cells = 3, min.features = 200)
dat6 <- CreateSeuratObject(counts = rawData6, project = "week6", min.cells = 3, min.features = 200)
dat3 <- CreateSeuratObject(counts = rawData3, project = "week3", min.cells = 3, min.features = 200)

datList <- list(datTu, dat9, dat6, dat3)
datList <- list(tumor = datTu, week9 = dat9, week6 = dat6, week3 = dat3)
#names(datList) <- c("tumor", "9weeks")

rm(datTu, dat9, dat6, dat3)

```

```{r}

sweepResList <- list()
sweepSumList <- list()
bcmvnList <- list()
homotypList <- list()
nExp_poi.adj <- list()
datListFilt <- list()

for (i in names(datList)) {
  datList[[i]][["qc.mito"]] <- PercentageFeatureSet(datList[[i]], pattern = "^mt-")
  
  print(VlnPlot(datList[[i]], features = c("nFeature_RNA", "nCount_RNA", "qc.mito"), ncol = 3))
  
  plot1 <- FeatureScatter(datList[[i]], feature1 = "nCount_RNA", feature2 = "qc.mito")
  plot2 <- FeatureScatter(datList[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
  
  print(plot1 + plot2)
  
  datList[[i]] <- subset(datList[[i]] , subset = nFeature_RNA > 200 & nCount_RNA > 1000 & nFeature_RNA < 6000
                               & qc.mito < 15)
  
  datList[[i]] <- NormalizeData(datList[[i]])
  datList[[i]] <- FindVariableFeatures(datList[[i]], selection.method = "vst", nfeatures = 2000)
  datList[[i]] <- ScaleData(datList[[i]])
  datList[[i]] <- RunPCA(datList[[i]])
  datList[[i]] <- RunUMAP(datList[[i]], dims = 1:20, reduction = "pca")
  datList[[i]] <- FindNeighbors(datList[[i]], dims = 1:20, reduction = "pca")
  datList[[i]] <- FindClusters(datList[[i]], resolution = c(0.5))
  
  sweepResList[[i]] <- paramSweep(datList[[i]], PCs = 1:20, sct = FALSE)
  sweepSumList[[i]] <- summarizeSweep(sweepResList[[i]], GT = FALSE)
  bcmvnList[[i]] <- find.pK(sweepSumList[[i]])
  homotypList[[i]] <- modelHomotypic(datList[[i]]@meta.data$seurat_clusters)
  nExp_poi <- round(0.075*nrow(datList[[i]]@meta.data))
  nExp_poi.adj[[i]] <- round(nExp_poi*(1-homotypList[[i]]))
  datList[[i]] <- doubletFinder(datList[[i]], PCs = 1:20, pN = 0.25, pK = max(bcmvnList[[i]]$BCmetric),
                              nExp = nExp_poi.adj[[i]], reuse.pANN = FALSE, sct = FALSE)
  seurat_colnames <- colnames(datList[[i]]@meta.data)
  seurat_colnames <- lapply(seurat_colnames, function(x) sub("_0.25.*", "", x))
  colnames(datList[[i]]@meta.data) <- seurat_colnames
  table(datList[[i]]@meta.data$DF.classifications)
  datListFilt[[i]] <- subset(datList[[i]], subset = DF.classifications == 'Singlet')
  
  #cat(paste("QC for time point", i))
}

```


### merge object 

```{r}

comboDat <- merge(x = datListFilt[["tumor"]], y = list(datListFilt[["week9"]], datListFilt[["week6"]], datListFilt[["week3"]]),
                  project = 'KrasMouse')

```

### running analysis prior to intergration 

```{r}

comboDat <- NormalizeData(comboDat)
comboDat <- FindVariableFeatures(comboDat)
comboDat <- ScaleData(comboDat)
comboDat <- RunPCA(comboDat)
ElbowPlot(comboDat, ndims = 40)

comboDat <- FindNeighbors(comboDat, dims = 1:30, reduction = "pca")
comboDat <- FindClusters(comboDat, resolution = c(0.1, 0.3, 0.5, 0.7, 1))

comboDat <- RunUMAP(comboDat, dims = 1:30, reduction = "pca")

```

# UMAP projections 

```{r}

umap01 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.0.1")
umap03 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.0.3")
umap05 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.0.5")
umap07 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.0.7")
umap1 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.1")
umapIdent <- DimPlot(comboDat, reduction = 'umap', group.by = "orig.ident")

grid.arrange(
  umap01, umap03, umap05, 
  umap07, umap1, umapIdent, 
  ncol = 3, nrow = 2,
  top = textGrob("Non-integrated conditions", gp=gpar(fontsize=20)) 
)

```


### intergreating layesr 

```{r}
comboDat <- IntegrateLayers(object = comboDat, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca",
    verbose = FALSE)

comboDat[["RNA"]] <- JoinLayers(comboDat[["RNA"]])

comboDat <- FindNeighbors(comboDat, reduction = "integrated.cca", dims = 1:30)
comboDat <- FindClusters(comboDat, resolution = c(0.1, 0.3, 0.5, 0.7, 1))

comboDat <- RunUMAP(comboDat, dims = 1:30, reduction = "integrated.cca")
#DimPlot(comboDat, reduction = "umap", group.by = c("orig.ident", "seurat_clusters"))

#table(Idents(comboDat))
```

# UMAP projections 

```{r}

umap01 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.0.1")
umap03 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.0.3")
umap05 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.0.5")
umap07 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.0.7")
umap1 <- DimPlot(comboDat, reduction = 'umap', group.by = "RNA_snn_res.1")
umapIdent <- DimPlot(comboDat, reduction = 'umap', group.by = "orig.ident")

grid.arrange(
  umap01, umap03, umap05, 
  umap07, umap1, umapIdent, 
  ncol = 3, nrow = 2,
  top = textGrob("Non-integrated conditions", gp=gpar(fontsize=20)) 
)

```


### investigatign different cluster resoloutions 

# exocrine cells 

```{r}
comboDat <- SetIdent(comboDat, value = "RNA_snn_res.0.5")
DotPlot(comboDat, features = c("Cpa1", "Amy1", "Amy2a2","Krt19","Krt18", "Krt8", "Sox9","Foxq1", "Onecut2", "Tff1")) + RotatedAxis()
comboDat <- RenameIdents(comboDat, "13" = "Ex1", 
                         "17" = "Ex2", 
                         "25" = "Krt and foxq", 
                         "28" = "Whats this", 
                         "12" = "High Cpa")
table(Idents(comboDat))
DimPlot(comboDat, reduction = 'umap', label = T)
```

```{r}
comboDat <- SetIdent(comboDat, value = "RNA_snn_res.0.7")
DotPlot(comboDat, features = c("Cpa1", "Amy1", "Amy2a2","Krt19","Krt18", "Krt8", "Sox9","Foxq1", "Onecut2", "Tff1")) + RotatedAxis()
comboDat <- RenameIdents(comboDat, "14" = "Ex1", 
                         "21" = "Ex2", 
                         "28" = "Krt and foxq", 
                         "32" = "Ex3", 
                         "31" = "Intrest", 
                         "13" = "High Cpa1")
table(Idents(comboDat))
DimPlot(comboDat, reduction = 'umap', label = T)
```

```{r}
comboDat <- SetIdent(comboDat, value = "RNA_snn_res.1")
DotPlot(comboDat, features = c("Cpa1", "Amy1", "Amy2a2","Krt19","Krt18", "Krt8", "Sox9","Foxq1", "Onecut2", "Tff1", "Pyy", "Sst", "Muc1")) + RotatedAxis()
VlnPlot(comboDat, features = c("Cpa1"))
VlnPlot(comboDat, features = c("Krt19"))
VlnPlot(comboDat, features = c("Foxq1"))
FeaturePlot(comboDat, features = c("Cpa1", "Krt19", "Foxq1", "Tff1"))

comboDat <- RenameIdents(comboDat, "14" = "Ex1", 
                         "24" = "Ex2", 
                         "32" = "Krt and foxq", 
                         "35" = "interest",
                         "36" = "Ex3")
table(Idents(comboDat))
DimPlot(comboDat, reduction = 'umap', label = T) + NoLegend()
```

# fibroblasts 

```{r}
VlnPlot(comboDat, features = c("Col1a1"))
DotPlot(comboDat, features = c("Col1a1"))
comboDat <- RenameIdents(comboDat, "1" = "Fibro1", 
                         "6" = "Fibro2", 
                         "8" = "Fibro3",
                         "18" = "Fibro4",
                         "31" = "Fibro5", 
                         "37" = "Fibro6")
table(Idents(comboDat))
DimPlot(comboDat, reduction = 'umap', label = T)
```

# other, inconclusive

```{r}
VlnPlot(comboDat, features = c("Resp18"))
DotPlot(comboDat, features = c("Resp18"))
VlnPlot(comboDat, features = c("Rgs5"))
DotPlot(comboDat, features = c("Rgs5"))
VlnPlot(comboDat, features = c("Pecam1"))
DotPlot(comboDat, features = c("Pecam1"))

DotPlot(comboDat, features = c("Pecam1", "Rgs5", "Resp18"))
FeaturePlot(comboDat, features = "Pecam1")

comboDat <- RenameIdents(comboDat, "13" = "Endothelium1", 
                         "33" = "Endothelium2", 
                         "34" = "Endothelium3")
table(Idents(comboDat))
DimPlot(comboDat, reduction = 'umap', label = T)
```

# immune cells 

```{r}
DotPlot(comboDat, features = c("Il7r", "Ccr7", "Ptprc", "Cd3d", "Cd3e", "Cd19", "Cd79b", "Amy1", "Amy2a2"))
comboDat <- RenameIdents(comboDat, "4" = "Tcell1", 
                         "5" = "Tcell2", 
                         "7" = "Tcell3", 
                         "10" = "Tcell4",
                         "17" = "Tcell5", 
                         "23" = "Tcell6", 
                         "33" = "Tcell and Bcell like", 
                         "0" = "Bcell1",
                         "12" = "Bcell",
                         "25" = "Bcell3",
                         "22" = "weak Tcell and Bcell like")
table(Idents(comboDat))
DimPlot(comboDat, reduction = 'umap', label = T) + NoLegend()
```

# macrophages

```{r}
DotPlot(comboDat, features = c("Cd14", "Adgre1", "Cd52"))
comboDat <- RenameIdents(comboDat, "2" = "Macrophage1", 
                         "3" = "Macrophage2", 
                         "9" = "Macrophage3", 
                         "19" = "Pos macrophage",
                         "28" = "Pos macrophage2")
table(Idents(comboDat))
DimPlot(comboDat, reduction = 'umap', label = T) + NoLegend()
```

# Endothelial 

```{r}
DotPlot(comboDat, features = c("Cdh5", "Pecam1"))
comboDat <- RenameIdents(comboDat, "13" = "Endothelium1", 
                         "30" = "Endothelium2", 
                         "34" = "Endothelium3")
table(Idents(comboDat))
DimPlot(comboDat, reduction = 'umap', label = T) + NoLegend()
```




### moving forwar with res 1, identifying markers 

# all markers

```{r}
allMarkers <- FindAllMarkers(comboDat)
allMarkers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1)

markersList <- list()
clusIds <- levels(allMarkers$cluster)

for (i in clusIds) {
  markersList[[i]] <- allMarkers[allMarkers$cluster == as.character(i), ]
}
```

```{r}
head(markersList[["15"]], 20) # "Foxp1", "Foxn3" 
head(markersList[["20"]], 20)
head(markersList[["11"]], 20)
head(markersList[["Ex3"]], 20)
head(markersList[["10"]], 20)
head(markersList[["2"]], 20)

DotPlot(comboDat, features = c("Foxp1", "Foxn3"))

markers15 <- markersList[["15"]]
markers15 <- markers15[markers15$p_val_adj < 0.01 & markers15$avg_log2FC > 2,]
markers20 <- markersList[["20"]]
markers11 <- markersList[["11"]]


```

```{r}
markers15_Ex <- FindMarkers(comboDat, ident.1 = 15, ident.2 = c("Ex1", "Ex2"))
head(markers15_Ex, n = 20)
```

### examining QC removed cells 

# doublets

```{r}
datDouble <- list()

for (i in names(datList)) {
  datDouble[[i]] <- subset(datList[[i]], subset = DF.classifications == 'Doublet')
}

comboDatDouble <- merge(x = datDouble[["tumor"]], y = list(datDouble[["week9"]], datDouble[["week6"]], datDouble[["week3"]]),
                  project = 'KrasMouse')
comboDatDouble <- NormalizeData(comboDatDouble)
comboDatDouble <- FindVariableFeatures(comboDatDouble)
comboDatDouble <- ScaleData(comboDatDouble)
comboDatDouble <- RunPCA(comboDatDouble)
#ElbowPlot(comboDat, ndims = 40)

comboDatDouble <- FindNeighbors(comboDatDouble, dims = 1:30, reduction = "pca")
comboDatDouble <- FindClusters(comboDatDouble, resolution = c(0.1, 0.3, 0.5, 0.7, 1))

comboDatDouble <- RunUMAP(comboDatDouble, dims = 1:30, reduction = "pca")


DimPlot(comboDatDouble, reduction = 'umap')
DotPlot(comboDatDouble, features = c("Cpa1", "Amy1", "Amy2a2","Krt19","Krt18", "Krt8", "Sox9","Foxq1", "Onecut2", "Tff1")) + RotatedAxis()
table(Idents(comboDatDouble))
```

# initial QC filters 

```{r}

datTu <- CreateSeuratObject(counts = rawDataTu, project = "tumor", min.cells = 3, min.features = 200)
dat9 <- CreateSeuratObject(counts = rawData9, project = "week9", min.cells = 3, min.features = 200)
dat6 <- CreateSeuratObject(counts = rawData6, project = "week6", min.cells = 3, min.features = 200)
dat3 <- CreateSeuratObject(counts = rawData3, project = "week3", min.cells = 3, min.features = 200)

QCremoveList <- list(datTu, dat9, dat6, dat3)
QCremoveList <- list(tumor = datTu, week9 = dat9, week6 = dat6, week3 = dat3)

QCremoveList <- list()

for (i in names(QCremoveList)) {
  
  QCremoveList[[i]][["qc.mito"]] <- PercentageFeatureSet(QCremoveList[[i]], pattern = "^mt-")

  
  QCremoveList[[i]] <- subset(QCremoveList[[i]] , subset = nFeature_RNA > 6000
                               & qc.mito > 15)
  
  QCremoveList[[i]] <- NormalizeData(QCremoveList[[i]])
  QCremoveList[[i]] <- FindVariableFeatures(QCremoveList[[i]], selection.method = "vst", nfeatures = 2000)
  QCremoveList[[i]] <- ScaleData(QCremoveList[[i]])
  QCremoveList[[i]] <- RunPCA(QCremoveList[[i]])
  QCremoveList[[i]] <- RunUMAP(QCremoveList[[i]], dims = 1:20, reduction = "pca")
  QCremoveList[[i]] <- FindNeighbors(QCremoveList[[i]], dims = 1:20, reduction = "pca")
  QCremoveList[[i]] <- FindClusters(QCremoveList[[i]], resolution = c(0.5))
  
  #cat(paste("QC for time point", i))
}

QCremoveList <- merge(x = QCremoveList[["tumor"]], y = list(QCremoveList[["week9"]], QCremoveList[["week6"]], QCremoveList[["week3"]]),
                  project = 'KrasMouse')
QCremoveList <- NormalizeData(QCremoveList)
QCremoveList <- FindVariableFeatures(QCremoveList)
QCremoveList <- ScaleData(QCremoveList)
comboDatDouble <- RunPCA(QCremoveList)
#ElbowPlot(comboDat, ndims = 40)

comboDatDouble <- FindNeighbors(comboDatDouble, dims = 1:30, reduction = "pca")
comboDatDouble <- FindClusters(comboDatDouble, resolution = c(0.1, 0.3, 0.5, 0.7, 1))

comboDatDouble <- RunUMAP(comboDatDouble, dims = 1:30, reduction = "pca")


DimPlot(comboDatDouble, reduction = 'umap')
DotPlot(comboDatDouble, features = c("Cpa1", "Amy1", "Amy2a2","Krt19","Krt18", "Krt8", "Sox9","Foxq1", "Onecut2", "Tff1")) + RotatedAxis()
table(Idents(comboDatDouble))
```














